Multiple Dice Working as One: CAD Flows and Routing Architectures for Silicon Interposer FPGAs

### 2014-12-09 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4529 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Re-adding aspect ratio figure
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4530 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
Vaughn and Javeed's meeting updates.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4531 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{Making Multiple Dice Work as One: CAD Flows and Routing Architectures for Silicon Interposer FPGAs}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

%   - Channel width vs % wires cut
\subsection{Interposer Wiring Capacity (\% wires cut)}

@@

%   - Impact of # of cutlines
\subsection{Impact of Number of Dice}
\label{num_dice_impact}
To examine the impact of the number of dice used to construct an interposer-based FPGA, we measured the minimum channel width for an FPGA device with four dice, varying the fraction of wires cut. We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the best architecture parameters from this section (TODO JAVEED -- spell them out?). The results, shown in Figure \ref{fig:wires_cut_4part}, show the same trend as Figure \ref{fig:wires_cut}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{wires_cut_4part.eps}
\caption{Impact of partitioning-based CAD flow on routability for a range of architectures with 4 dice. TODO JAVEED -- LU32 90\% result is missing}
\label{fig:wires_cut_4part}
\end{figure}

TODO JAVEED -- show critical path delay? (it's uninteresting)

%   - Impact of aspect ratio
\subsection{Impact of Aspect Ratio}
\label{aspect_ratio_impact}
TODO: Javeed
We investigated the implications of the device aspect ratio on routability for the four-partition case. In particular, we compared a device with a square interposer (four rectangular dice) to a device with a rectangular interposer (four square dice), as shown in \ref{fig:square_tall}. We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the best architecture parameters from this section. The results are shown in Table \ref{table:squares}.

@@

\begin{table}[!htbp]
\centering
\begin{tabular}{|l|r|r|}
\hline
 & Square Interposer & Square Dice \\ \hline \hline
bgm & 102 & 306 \\ \hline
stereovision0 & 86 & 256 \\ \hline
stereovision1 & 166 & 256 \\ \hline
LU8PEEng & 156 & 126 \\ \hline
mkDelayWorker32B & 96 & 194 \\ \hline
mcml & 196 & 496 \\ \hline
LU32PEEng & 256 & 264 \\ \hline
geomean & 140.7481367 & 251.805728 \\ \hline
\end{tabular}
\caption{Minimum channel width for square interposer and square dice configurations.}
\label{table:squares}
\end{table}

The square interposer configuration performs significantly better than the square dice configuration.

TODO JAVEED -- Vaughn commented that we should talk about which is easier to manufacture. I think the square dice device would actually be easier to manufacture, because the more complex dice will fit well on a round wafer. Isn't it fine for the interposer to be non-square?

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{Multiple Dice Working as One: CAD Flows and Routing Architectures for Silicon Interposer FPGAs}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

%   - Channel width vs % wires cut
\subsection{Interposer Wiring Capacity (\% wires cut)}
\label{sec:resultsWiringCapacity}

@@

%   - Impact of # of cutlines
\subsection{Impact of Number of Dice}
\label{num_dice_impact}
To examine the impact of the number of dice used to construct an interposer-based FPGA, we compare the minimum channel width for FPGA systems composed of four dice and two dice, respectively, while varying the fraction of wires cut at the interposer boundaries. We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the same interposer routing architecture parameters as Section \ref{sec:resultsWiringCapacity}. 
As Figure \ref{fig:wires_cut_4part} shows, four-die systems perform almost as well in routability as two-die systems until the \textit{\% wires cut} exceeds 80\%, at which point routability is worse for the four-die system. 
TODO JAVEED -- add extra line to graph, make sure same circuits for all points.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{wires_cut_4part.eps}
\caption{Impact of number of dice on routability. TODO JAVEED -- LU32 90\% result is missing}
\label{fig:wires_cut_4part}
\end{figure}

Systems with more dice are slower, on average the critical path delay is xx\% higher for a four-die system than a two-die system. This speed difference is not significantly affected by the \textit{\% wires cut}.
 TODO JAVEED -- figure out critical path delay %.

%   - Impact of aspect ratio
\subsection{Impact of Aspect Ratio}
\label{aspect_ratio_impact}
It is easier to manufacture silicon dice that have an aspect ratio near 1 (i.e. square). Commercial FPGAs are laid out with identical columns, and hence combine multiple identical FPGA dice in \textit{one dimension} on an interposer \cite{xilinxTSV}. As Figure \ref{fig:square_tall} shows, we can make \textit{either} the FPGA dice square, or the interposer square, but not both. Table \ref{table:squares} compares the routability of these two options for the four-die case: a square interposer (four rectangular dice) or a rectangular interposer (four square dice.) We used the best CAD flow from Section \ref{sec:partitioner_stage_results} and the same routing architecture parameters as Section \ref{sec:resultsWiringCapacity}.

@@

\begin{table}[!htbp]
\centering
\begin{tabular}{|l|r|r|}
\hline
 & Square Interposer & Square Dice \\ \hline \hline
bgm & 102 & 306 \\ \hline
stereovision0 & 86 & 256 \\ \hline
stereovision1 & 166 & 256 \\ \hline
LU8PEEng & 156 & 126 \\ \hline
mkDelayWorker32B & 96 & 194 \\ \hline
mcml & 196 & 496 \\ \hline
LU32PEEng & 256 & 264 \\ \hline
geomean & 141 & 252 \\ \hline
\end{tabular}
\caption{Minimum channel width for square interposer and square dice configurations.}
\label{table:squares}
\end{table}

The square interposer configuration performs significantly better than the square dice configuration. A square interposer results in cutlines that are closer together, and hence more logic blocks are near a cutline. On the other hand, a square interposer (composed of rectangular FPGA dice) has more vertical routing channels, giving it a higher absolute number of wires between FPGA dice. The second effect is the more important one, leading to much better routability for a square interposer.
